---
title: "Final Project_Team 7_Gabi"
author: "Gabi Rivera Liam Richardson Sahil Wadhwa"
format: pdf
editor: visual
---

# **Sensor-Fusion Smoke Detection Classification**

The goal is to devise a Machine Learning model that that will detect smoke through the use of loT data to trigger a fire alarm.

**Information about the Features:**

*Air Temperature*\
*Air Humidity*\
*TVOC:* Total Volatile Organic Compounds; measured in parts per billion (Source)\
*eCO2:* co2 equivalent concentration; calculated from different values like TVCO\
*Raw H2:* raw molecular hydrogen; not compensated (Bias, temperature, etc.)\
*Raw Ethanol:* raw ethanol gas (Source)\
*Air Pressure*\
*PM 1.0 and PM 2.5:* particulate matter size \< 1.0 µm (PM1.0). 1.0 µm \< 2.5 µm (PM2.5)\
*Fire Alarm:* ground truth is "1" if a fire is there\
*CNT:* Sample counter\
*UTC:* Timestamp UTC seconds\
*NC0.5/NC1.0 and NC2.5:* Number concentration of particulate matter. This differs from PM because NC gives the actual number of particles in the air. The raw NC is also classified by the particle size: \< 0.5 µm (NC0.5); 0.5 µm \< 1.0 µm (NC1.0); 1.0 µm \< 2.5 µm (NC2.5);

## Pre-Processing:

```{r warning=FALSE, message=FALSE}
# List of libraries
library(caret)
library(tidyverse)
library(naniar)
library(gt)
library(ggplot2)
library(dplyr)
library(tidyr)
library(GGally)
library(corrplot)
library(e1071)
library(tibble)
library(MASS)
library(mice)
library(reshape2)
library(ROSE)
library(pROC)
library(lubridate)
```

Note:

-   Train dataset has 5000 observations, 14 predictors, and an outcome variable.
-   Test dataset has 12437 observations and 14 predictors.

```{r dataset}
# Upload datasets
train <- read.csv("train_dataset.csv")
test <- read.csv("test_dataset.csv")

# Use train_dataset moving forward
smokedf <- train 
str(smokedf)
class(smokedf)
#summary(smokedf)
```

```{r missing}
# Missing Values
smokedf[smokedf == ""] <- NA
na_value <- sapply(smokedf, function(x) sum(is.na(x)))
predictors_with_missing <- names(na_value[na_value > 0])

missing_values_table <- data.frame(Predictor = predictors_with_missing,
Missing_Values = na_value[predictors_with_missing])
missing_values_table |> head() |> gt() |>
  tab_header(title = "Predictors with Missing Values")

vis_miss(smokedf)
```

```{r histogram}
# Histogram of each predictors
smokedf |>
  pivot_longer(-Fire.Alarm, names_to = 'Element', values_to = 'value') |>
  ggplot(aes(x = value)) +
  geom_histogram(bins = 20) +
  facet_wrap(~Element, scales = 'free', ncol = 4) +  
  theme(axis.text.x = element_text(angle = 10)) +
  labs(title = 'Predictors Distribution', x = NULL, y = "Count")
```

```{r box plot I}
# Create box plots for each predictors:
smokedf |>
  pivot_longer(-Fire.Alarm, names_to = 'Element', values_to = 'value') |>
  ggplot(aes(value)) +
  geom_boxplot() +
  facet_wrap(~Element, scales = "free", ncol = 4) +
  theme(axis.text.x = element_text(angle = 10)) +
  labs(title = 'Box Plots of Predictors', x = "Count")
```

```{r box plot II}
# Create box plots of response for each predictors:
smokedf |>
  pivot_longer(-Fire.Alarm, names_to = 'Element', values_to = 'value') |>
  ggplot(aes(Fire.Alarm, value, group = Fire.Alarm)) +
  geom_boxplot() +
  facet_wrap(~Element, scales = "free", ncol = 4) +
  theme(axis.text.x = element_text(angle = 0)) +
  labs(title = 'Box Plots of Predictors over Response', x = "Fire Alarm", y = "Count")
```

```{r UTC}
# Convert UTC format into hours of day and day of week
smokedf$UTC_datetime <- as.POSIXct(smokedf$UTC, origin = "1970-01-01", tz = "UTC")
# all from 2022 and June

smokedf$hour <- hour(smokedf$UTC_datetime)
smokedf$day <- wday(smokedf$UTC_datetime, week_start = 1)

# Reorder variables and remove UTC_datetime
other_columns <- setdiff(names(smokedf), "Fire.Alarm")
smokedf <- smokedf[, c(other_columns, "Fire.Alarm")]
smokedf <- subset(smokedf, select = c(-UTC_datetime, -UTC))
```

```{r correlation}
# Correlation plot including outcome variable
smokedf |>
mutate(Fire.Alarm = as.numeric(Fire.Alarm)) |>
cor() |>
corrplot.mixed(title = "Correlation Plot Between Outcome and Predictors",
tl.cex = .4, number.cex = 0.6, mar = c(1, 1, 1, 1))
```

```{r reduced}
# Remove highly correlated predictors
correlation_matrix <- cor(smokedf[, -1])
highly_correlated <- findCorrelation(correlation_matrix, cutoff = 0.75)
highly_correlated_names <- colnames(smokedf[, -1])[highly_correlated]
smokedf_reduced <- smokedf[, -highly_correlated]

# Correlation plot including outcome variable
smokedf_reduced |> 
  mutate(Fire.Alarm = as.numeric(Fire.Alarm)) |>
  cor() |>
  corrplot.mixed(title = "Reduced Correlation Plot",
                 tl.cex = .5, number.cex = 0.8, mar = c(1, 1, 1, 1))
```

```{r skew, warning=FALSE, message=FALSE}
# Calculate skewness of each predictors
# Skew values less than ±0.5 should be considered ‘normal enough’
skew_fa <- smokedf_reduced |>
  dplyr::select(-Fire.Alarm) |>
  map_dbl(skewness) |> round(3) 

skew_fa_tibble <- tibble(Variable = names(skew_fa), Skewness = skew_fa)
skew_fa_tibble |> gt()

# Determine zero values in the dataframe
zero_counts <- sapply(smokedf, function(x) sum(x == 0, na.rm = TRUE))
```

```{r box cox}
# Perform Box Cox Tranformation Analysis and determine predictors with >50% improvement
bct_test <- function(x, property = 'skew') {
  stopifnot(property %in% c('skew', 'lambda'))
  
  x <- x[which(! is.na(x))] 
  x2 <- x + ifelse(any(x == 0), 0.0001, 0) 

  bct <- BoxCoxTrans(x2)
  x_trans <- predict(bct, x2 )
if (property == 'skew') return(e1071::skewness(x_trans))
return(bct$lambda)}

skew_smfa_bct <- smokedf_reduced |>
  dplyr::select(-Fire.Alarm) |>
  map_dbl(bct_test)

bct_analysis <- tibble(
  Property = names(skew_fa),
  `Original Skew` = skew_fa,
  `Skew after BoxCox`= round(skew_smfa_bct,4),
  Lambda = smokedf_reduced |> dplyr::select(-Fire.Alarm) |> 
    map_dbl( ~ bct_test(.x, 'lambda')))

bct_keep <- bct_analysis |>
  filter(abs(`Original Skew`) > 0.5,
         abs(`Skew after BoxCox`) < 0.5)

bct_keep |>
gt::gt()

# Apply BCT Analysis result in reduced dataframe
bct <- BoxCoxTrans(smokedf_reduced$CNT + 
                     ifelse(any(smokedf_reduced$CNT == 0), 0.0001, 0))
smokedf_reduced$gamma_CNT <- predict(bct, smokedf_reduced$CNT)

# Plot Changes
smokedf_reduced |> ggplot(aes(x = gamma_CNT)) +
    geom_histogram(binwidth = 500, fill = "pink", color = "white") +
    geom_histogram(aes(x = CNT), binwidth = 500, fill = "darkgray", alpha = 10) +
    labs(title = "Histogram of Box-Cox Transformed gamma with 0")

smokedf_reduced |>
ggplot(aes(x = gamma_CNT)) +
geom_histogram(bins= 9, color = "white", fill = "pink") +
labs(title ="Histogram of Box-Cox Transformed gamma_CNT")

# Remove Original CNT in reduced dataframe
smokedf_reduced <- smokedf_reduced |> dplyr::select(-CNT) |> 
  relocate("Fire.Alarm", .after = last_col())
```

## Exploratory Data Analysis

```{r orginal, warning=FALSE, message=FALSE}
# General relationship plot: Multivariate Analysis (Original Data)
smokedf |>
  dplyr::select(-Fire.Alarm) |>
  ggpairs(title = 'Original: Predictors Relationship Map', progress = TRUE,
          upper = list(continuous = wrap("cor", size = 2))) +
  theme_grey(base_size = 5) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 5),
        axis.text.y = element_text(size = 5),
        panel.spacing = unit(0.01, "lines"))
```

```{r reduced data, warning=FALSE, message=FALSE}
# General relationship plot: Multivariate Analysis (Reduced Data)
smokedf_reduced$Fire.Alarm <- as.factor(smokedf_reduced$Fire.Alarm)

smokedf_reduced |>
  ggpairs(title = 'Reduced: Predictors Relationship Map', progress = TRUE,
          upper = list(continuous = wrap("cor", size = 2.5)),
          lower = list(continuous = wrap("points", alpha = 0.3), 
                       combo = wrap("box_no_facet", alpha = 0.4)),
          columns = 1:9, ggplot2::aes(colour = Fire.Alarm)) +
  theme_grey(base_size = 8) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 5),
        axis.text.y = element_text(size = 8),
        panel.spacing = unit(0.1, "lines"))
```

```{r proportion}
# Explore proportion of Fire.Alarm classes against each predictors
proportions <- smokedf_reduced |>
  pivot_longer(-Fire.Alarm, names_to = 'variable', values_to = 'value') |>
  group_by(Fire.Alarm, variable, value) |>
  summarize(Count = n(), .groups = 'drop') |>
  group_by(variable, Fire.Alarm) |>
  mutate(Proportion = Count / sum(Count))

# Create the bar plot
proportions |>
  ggplot(aes(x = Fire.Alarm, y = Proportion, fill = Fire.Alarm)) +
  geom_bar(stat = "identity", position = "dodge") +
  facet_wrap(~ variable, scales = "free_y", ncol = 4) +
  labs(x = "Fire.Alarm", y = "Proportion") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 5))
```

```{r range}
# Explore the the value range of Fire.Alarm classes against each predictors
smokedf_reduced |>
  pivot_longer(-Fire.Alarm, names_to = 'variable', values_to = 'value') |>
  ggplot(aes(x = Fire.Alarm, y = value, fill = Fire.Alarm)) +
  geom_bar(stat = "identity", position = "dodge") +
  facet_wrap(~ variable, scales = "free_y", ncol = 4) +
  labs(x = "Fire.Alarm", y = "Range of Value") +
  theme_minimal()
```

```{r hours}
# Prevelent hour of the day
smokedf$Fire.Alarm1 <- factor(smokedf$Fire.Alarm, levels = c(0, 1), labels = c("No", "Yes"))

smokedf|>
  ggplot(aes(x = factor(hour), fill = Fire.Alarm1)) +
  geom_bar(position = "stack") +
  labs(title = 'Frequency of Fire Alarm Classes Per Hour', x = "Hour of Day", 
       y = "Count", fill = "Fire Alarm") +
  scale_x_discrete(labels = function(x) as.numeric(x)) +
  scale_fill_manual(values = c("Yes" = "pink", "No" = "darkgray")) +  
  theme_minimal()
```

```{r week}
# Prevelent day of the week
smokedf |>
  ggplot(aes(x = factor(day), fill = Fire.Alarm1)) +
  geom_bar(position = "stack") +
  labs(title = 'Frequency of Fire Alarm Classes Per Day', x = "Day of Week", 
       y = "Count", fill = "Fire Alarm") +
  scale_x_discrete(labels = function(x) as.numeric(x)) +
  scale_fill_manual(values = c("Yes" = "darkgreen", "No" = "darkgray")) +  
  theme_minimal()
```

## Training and Test Sets

```{r tt}
set.seed(503)
smokedf_reduced$Fire.Alarm <- ifelse(smokedf_reduced$Fire.Alarm == 0, "No", "Yes")
smokedf_predictors <- smokedf_reduced[ ,-1]
smokedf_yield <- smokedf_reduced[ ,1]

# Split the data into training and test sets
smokedf_index <- createDataPartition(smokedf_yield, p = 0.8, list = FALSE)
smokedf_train <- smokedf_predictors[smokedf_index, ]
smokedf_test <- smokedf_predictors[-smokedf_index, ]
```

```{r csb}
# Center and scaling using preProcess
smokedf_prep <- preProcess(smokedf_train, method = c("center", "scale"))

# Apply the transformation to the training data
smokedf_train_transformed <- predict(smokedf_prep, smokedf_train)
smokedf_test_transformed <- predict(smokedf_prep, smokedf_test)

# Rebalance train dataset
cat("Number of non-triggered fire alarm (No) = ", 
    sum(smokedf_train_transformed$Fire.Alarm == "No"), "\n") 
cat("Number of triggered fire alarm (Yes) =", 
    sum(smokedf_train_transformed$Fire.Alarm == "Yes"))

train_balanced <- ROSE(Fire.Alarm ~ ., data = smokedf_train_transformed)$data

rebalanced <- as.data.frame(table(train_balanced$Fire.Alarm)) 
rebalanced |> gt() |> 
  tab_header(title = "Rebalanced Fire Alarm Counts")
```

## Model Strategies

```{r lr}
set.seed(503)
ctrl <- trainControl(method = "cv", summaryFunction = twoClassSummary, 
                     classProbs = TRUE, savePredictions = TRUE)

# Create Logistic Regression
lrFit <- train(Fire.Alarm ~ ., 
               data = train_balanced, 
               method = "glm", 
               metric = "ROC", 
               trControl = ctrl)
# Logistic regression final model using train data set
lrFit$finalModel
lrFit

# Confusion matrix
lrCM <- confusionMatrix(lrFit, norm = "none")
lrCM

# Logistic Regression ROC curve and AUC score
lrRoc <- roc(response = lrFit$pred$obs,
             predictor = lrFit$pred$Yes,
             levels = rev(levels(lrFit$pred$obs)))
auc_score_lr <- auc(lrRoc )

plot(lrRoc, main = "Logistric Regression Train Model ROC Curve", 
     col = "darkgreen", lwd = 2)
abline(a = 0, b = 1, lty = 3, col = "darkgray")
text(0.5, 0.3, paste0("AUC = ", round(auc_score_lr, 2)), adj = 2.6, col = "darkgreen")
legend("bottomright", legend = c("LR ROC Curve", "Random"), 
       col = c("darkgreen", "darkgray"), lty = c(1, 2), lwd = c(2, 1))

# Variable importance
lrImp <- varImp(lrFit, scale = FALSE)
plot(lrImp, main = "Variable Importance (Logistic Regression)",
     xlab = "Importance (Std.Dev. of Coefficients)", ylab = "Variables")
```

```{r gmb}
# Create Penalized Logistic Regression
glmnGrid <- expand.grid(alpha = seq(0, 1, by = 0.1),
lambda = seq(.001, .3, length = 10))
glmnFit <- train(Fire.Alarm ~ ., 
                 data = train_balanced, 
                 method = "glmnet",
                 tuneGrid = glmnGrid,
                 preProc = c("center", "scale"),
                 metric = "ROC",
                 trControl = ctrl)
optimal_a <- glmnFit$bestTune$alpha
optimal_l <- glmnFit$bestTune$lambda
glmnmodel <- train(Fire.Alarm ~ .,
                   data = train_balanced,
                   method = "glmnet",
                   preProc = c("center", "scale"),
                   metric = "ROC",
                   trControl = ctrl,
                   tuneGrid = expand.grid(alpha = optimal_a,
                                          lambda = optimal_l))

glmnmodel
plot(glmnFit)

# Confusion matrix
glmnCM <- confusionMatrix(glmnmodel, norm = "none")
glmnCM

# Penalized Logistic Regression ROC curve and AUC score
glmnRoc <- roc(response = glmnmodel$pred$obs,
             predictor = glmnmodel$pred$Yes,
             levels = rev(levels(glmnmodel$pred$obs)))
auc_score_glmn <- auc(glmnRoc )

plot(glmnRoc, main = "Penalized Logistric Regression Train Model ROC Curve", 
     col = "pink", lwd = 2)
abline(a = 0, b = 1, lty = 3, col = "darkgray")
text(0.5, 0.3, paste0("AUC = ", round(auc_score_glmn, 2)), adj = 2.6, col = "pink")
legend("bottomright", legend = c("GLMNet ROC Curve", "Random"), 
       col = c("pink", "darkgray"), lty = c(1, 2), lwd = c(2, 1))
                    
# Variable importance
glmnImp <- varImp(glmnmodel, scale = FALSE)
plot(glmnImp, main = "Variable Importance (Penalized Logistic Regression)",
     xlab = "Importance (Std.Dev. of Coefficients)", ylab = "Variables")
```

```{r nsc}
# Create Nearest Shrunken Centroids
nscGrid <- expand.grid(threshold = seq(0, 25, length = 30))
nscFit <- train(Fire.Alarm ~ ., 
                data = train_balanced,
                method = "pam",
                tuneGrid = nscGrid,
                metric = "ROC",
                trControl = ctrl)
nscFit
plot(nscFit)

# Confusion matrix
nscCM <- confusionMatrix(nscFit, norm = "none")
nscCM

# Nearest Shrunken Centroids ROC curve and AUC score
nscRoc <- roc(response = nscFit$pred$obs,
             predictor = nscFit$pred$Yes,
             levels = rev(levels(nscFit$pred$obs)))
auc_score_nsc <- auc(nscRoc)

plot(nscRoc, main = "Nearest Shrunken Centroids Train Model ROC Curve", 
     col = "darkblue", lwd = 2)
abline(a = 0, b = 1, lty = 3, col = "darkgray")
text(0.5, 0.3, paste0("AUC = ", round(auc_score_glmn, 2)), adj = 2.6, col = "darkblue")
legend("bottomright", legend = c("NSC ROC Curve", "Random"), 
       col = c("darkblue", "darkgray"), lty = c(1, 2), lwd = c(2, 1))
                    
# Variable importance
nscImp <- varImp(nscFit, scale = FALSE)
plot(nscImp, main = "Variable Importance (Nearest Shrunken Centroids Regression)",
     xlab = "Importance (Std.Dev. of Coefficients)", ylab = "Variables")
```

```{r KNN}

set.seed(100)
knnTune <- train(Fire.Alarm ~ ., data = train_balanced,
                method = "knn",
                preProc = c("center","scale"),
                tuneGrid = data.frame(k = 1:20))

knnTune
plot(knnTune)

```

```{r MARS}

set.seed(100)
marsTune <- train(Fire.Alarm ~ ., data = train_balanced,
                  method = "earth",
                  tuneGrid = expand.grid(degree = 1, nprune = 2:38))
marsTune
plot(marsTune)

```

```{r SVM}

```

```{r NN}
# X_test <- df <- subset(test_df, select = -c(UTC, CNT, Fire.Alarm))

df <- read.csv("train_dataset.csv")  # Make sure to set the correct path
y_failure_type <- factor(df$Fire.Alarm)
# Remove unnecessary columns
X <-  subset(df, select = -c(UTC, CNT, Fire.Alarm))

if(any(!sapply(X, is.numeric))) {
  X[] <- lapply(X, function(x) as.numeric(as.factor(x)))
}

# Standardize numeric features
preProcValue <- preProcess(X,method = c("center", "scale"))
X_processed <- predict(preProcValue, X)

# Splitting the data into training and testing sets
set.seed(42)
trainIndex <- createDataPartition(y_failure_type, p = 0.7, list = FALSE)
X_train <- X_processed[trainIndex, ]
X_test  <- X_processed[-trainIndex, ]
y_train <- y_failure_type[trainIndex]
y_test  <- y_failure_type[-trainIndex]

# Convert data to torch tensors
X_train_tensor <- torch_tensor(as.matrix(X_train), dtype = torch_float())
X_test_tensor <- torch_tensor(as.matrix(X_test), dtype = torch_float())
y_train_tensor <- torch_tensor(as.integer(y_train), dtype = torch_long())
y_test_tensor <- torch_tensor(as.integer(y_test), dtype = torch_long())

class_counts <- table(y_train)
total_counts <- sum(class_counts)
manual_class_weights <- c(1.0, 3.0)  # Adjust these values as needed
class_weights <- torch_tensor(manual_class_weights, dtype = torch_float())

# Define the neural network class
Net <- nn_module(
  initialize = function(input_size, num_classes = 2) {
    self$fc1 <- nn_linear(input_size, 8)
    self$relu <- nn_relu()
    self$fc2 <- nn_linear(8, num_classes)
  },
  forward = function(x) {
    x <- self$fc1(x)
    x <- self$relu(x)
    x <- self$fc2(x)
    x
  }
)




```

```{r}
input_size <- ncol(X_train)

# Create an instance of the Net class
nn_model <- Net(input_size)

# Define a loss function and an optimizer
criterion <- nn_cross_entropy_loss(weight = class_weights)
optimizer <- optim_sgd(nn_model$parameters, lr = 0.01)

# Training loop
num_epochs <- 850000
for (epoch in 1:num_epochs) {
  nn_model$train()
  
  # Forward pass
  outputs <- nn_model(X_train_tensor)
  loss <- criterion(outputs, y_train_tensor)
  
  # Backward and optimize
  optimizer$zero_grad()
  loss$backward()
  optimizer$step()
  
  if (epoch %% 10000 == 0) {
    cat(sprintf("Epoch [%d/%d], Loss: %.4f\n", epoch, num_epochs, loss$item()))
  }
}

# Evaluation on the test set
# Evaluation on the test set


```

SVM

```{r}

svm_tune_grid=expand.grid(C = 2^(-5:2), sigma = 2^(-5:2))
svm_model <- train(X_train, y_train, method = 'svmRadial',
                   trControl = trainControl(method = 'cv', number = 10),
                   tuneGrid = svm_tune_grid,
                   tuneLength = 10)


svm_predictions <- predict(svm_model, X_test)

```

## Validation and Testing

```{r lrII}
# Predict test data using logistic regression
test_predictors <- smokedf_test_transformed[, -8]
test_org <- data.frame(Fire.Alarm = smokedf_test_transformed$Fire.Alarm)
test_org$Fire.Alarm <- factor(test_org$Fire.Alarm)

lr_predictions <- predict(lrFit, newdata = test_predictors, type = "raw")
lr_predictions <- data.frame(Fire.Alarm = lr_predictions)

# Confusion Matrix
lrCM_test <- confusionMatrix(data = as.factor(lr_predictions$Fire.Alarm), 
                             reference = test_org$Fire.Alarm, positive = "Yes")
lrCM_test
```

```{r glmnII}
# Predict test data using penalized logistic regression
glmn_predictions <- predict(glmnmodel, newdata = test_predictors, type = "raw")
glmn_predictions <- data.frame(Fire.Alarm = glmn_predictions)

# Confusion Matrix
glmnCM_test <- confusionMatrix(data = as.factor(glmn_predictions$Fire.Alarm),
                               reference = test_org$Fire.Alarm, positive = "Yes")
glmnCM_test
```

```{r nscII}
# Predict test data using nearest shrunken centroid
nsc_predictions <- predict(nscFit, newdata = test_predictors, type = "raw")
nsc_predictions <- data.frame(Fire.Alarm = nsc_predictions)

# Confusion Matrix
nscCM_test <- confusionMatrix(data = as.factor(nsc_predictions$Fire.Alarm),
                               reference = test_org$Fire.Alarm, positive = "Yes")
nscCM_test
```

```{r knn}

#Predictors and Confusion Matrix for KNN

test_predictors <- smokedf_test_transformed[, -8]
test_org <- data.frame(Fire.Alarm = smokedf_test_transformed$Fire.Alarm)
test_org$Fire.Alarm <- factor(test_org$Fire.Alarm)

knn_predictions <- predict(knnTune, newdata = test_predictors, type = "raw")
knn_predictions <- data.frame(Fire.Alarm = knn_predictions)

cm_knn <- confusionMatrix(data = as.factor(knn_predictions$Fire.Alarm),
                               reference = test_org$Fire.Alarm, positive = "Yes")
cm_knn

```

```{r mars}

#Predictors and Confusion Matrix for Mars

mars_predictions <- predict(marsTune, newdata = test_predictors, type = "raw")
mars_predictions <- data.frame(Fire.Alarm = mars_predictions)

cm_mars <- confusionMatrix(data = as.factor(mars_predictions$Fire.Alarm),
                               reference = test_org$Fire.Alarm, positive = "Yes")
cm_mars

```

```{r}

```

```{r}

```

## Performance Evaluation

```{r lr glmnet}
roc_test_lr <- roc(response = as.factor(test_org$Fire.Alarm),
                 predictor = as.numeric(lr_predictions$Fire.Alarm))
roc_test_glmn <- roc(response = as.factor(test_org$Fire.Alarm),
                 predictor = as.numeric(glmn_predictions$Fire.Alarm))
roc_test_nsc <- roc(response = as.factor(test_org$Fire.Alarm),
                 predictor = as.numeric(nsc_predictions$Fire.Alarm))

auc_test_lr <- auc(roc_test_lr)
auc_test_glmn <- auc(roc_test_glmn)
auc_test_nsc <- auc(roc_test_nsc)

# Plot ROC curve
plot(roc_test_lr, col = "darkgreen", main = "ROC Curves Comparison Using Test Data", lwd = 2)
plot(roc_test_glmn, col = "pink", add = TRUE, lwd = 2)
plot(roc_test_nsc, col = "darkblue", add = TRUE, lwd = 2)
legend("bottomright", legend = c(paste("LR Model (AUC =", round(auc_test_lr, 2), ")"),
                                 paste("GLMN Model (AUC =", round(auc_test_glmn, 2), ")"),
                                 paste("NSC Model (AUC =", round(auc_test_nsc, 2), ")")),
       col = c("darkgreen", "pink", "darkblue"), lwd = 2)

```

```{r compare}
# Model's ROC score
lrFit_ROC <- round(lrFit$results$ROC,3)
glmnmodel_ROC <- round(glmnmodel$results$ROC,3)
nscFit_ROC <- round(nscFit$results$ROC,3)

# Confusion matrix train accuracy score
lrCM_train_accuracy <- 0.924
glmnCM_train_accuracy <- 0.926
nscCM_train_accuracy <- 0.930

# Confusion matrix test AUC score
lrCM_AUC <- round(auc(roc_test_lr),3)
glmnCM_AUC <- round(auc(roc_test_glmn),3)
nscCM_AUC <- round(auc(roc_test_nsc),3)

# Confusion matrix test accuracy score
lrCM_test_accuracy <- round(lrCM_test$overall["Accuracy"],3)
glmnCM_test_accuracy <- round(glmnCM_test$overall["Accuracy"],3)
nscCM_test_accuracy <- round(nscCM_test$overall["Accuracy"],3)

# Confusion matric test class balanced accuracy score
lrCM_test_bal_accuracy <- round(lrCM_test$byClass["Balanced Accuracy"],3)
glmnCM_test_bal_accuracy <- round(glmnCM_test$byClass["Balanced Accuracy"],3)
nscCM_test_bal_accuracy <- round(nscCM_test$byClass["Balanced Accuracy"],3)

# Adjust nscFit_ROC
nscFit_ROC_adjusted <- nscFit_ROC[1]

# Now, create the data frame with adjusted nscFit_ROC
results_table <- data.frame(
  Model = c("LR", "Penalized LR", 
            "NSC"),
  ROC_Score = c(lrFit_ROC, glmnmodel_ROC, nscFit_ROC_adjusted),
  Train_Accuracy = c(lrCM_train_accuracy, glmnCM_train_accuracy, 
                     nscCM_train_accuracy),
  Test_AUC = c(lrCM_AUC, glmnCM_AUC, nscCM_AUC),
  Test_Accuracy = c(lrCM_test_accuracy, glmnCM_test_accuracy, 
                    nscCM_test_accuracy),
  Test_Balanced_Accuracy = c(lrCM_test_bal_accuracy, 
                             glmnCM_test_bal_accuracy, 
                             nscCM_test_bal_accuracy))

# Create a gt table
results_table |> gt() |>
  tab_header(title = "Model Comparison",
             subtitle = "Comparison of performance metrics for different models") |>
  cols_label(Model = "Model",
             ROC_Score = "Model ROC",
             Train_Accuracy = "Train Accuracy",
             Test_AUC = "Test AUC",
             Test_Accuracy = "Test Accuracy",
             Test_Balanced_Accuracy = "Test Balanced Accuracy")
```

```{r knn}

#Accuracy of 0.835

```

```{r mars}

#Accuracy of 0.757

```

NNET Eval

```{r}
nn_model$eval()
outputs_test <- nn_model(X_test_tensor)
predicted <- outputs_test$argmax(dim = 2)$sub(1L)  # Adjust predictions to match factor levels (1-based indexing)

predicted <- as.array(predicted)
predicted <- as.integer(predicted)

# Calculate accuracy and other metrics using confusionMatrix
nnet_confusion <- confusionMatrix(factor(predicted, levels = levels(y_test)), y_test)
print(nnet_confusion)
```

SVM Eval

```{r}
confusion_svm <- confusionMatrix(predictions, y_test)
print(confusion_svm)
```
